---
title: "dataVis"
author: "Sophie Castel"
date: "7/18/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Ideas
\begin{itemize}
\item Stack graphs on top of one another, calculate the delta between interpolated point and original point, colour it, stack the lengths like a histogram for each interpolation technique. 
\item TWO SECTIONS: 1. How the HW performs in light of perturbations to MT TT and WT 2. How the HW stacks up against the others

\end{itemize}


```{r}
library(ggplot2)
library(plotly)
```
Let's start with just a trivial visualization. Colouring in the area between lines representing error using \verb+polygon+

```{r}
N <- length(IntData[[1]][[1]][[1]][[1]])
D <- length(IntData)
error <- numeric(N)
meanError <- numeric(N)
D_vec <- numeric(D)
p <- numeric(D)

for(d in 1:length(IntData)){
for(k in 1:N){
  
  intTest <- IntData[[d]][[1]][[4]][[3]][[k]] #d1,m18,p0.05,g1
  orgTest <- OriginalData[[d]]

p <- ggplot() +  
  geom_ribbon(aes(ymin = (as.numeric(intTest)), ymax = (as.numeric(orgTest)), x = 0:(length(intTest)-1)), fill = "mistyrose2") +
  
  geom_line(data = orgTest, aes(x = 0:(length(orgTest)-1), y = as.numeric(orgTest)), color = "lightpink3") +  
  geom_line(data = intTest, aes(x = 0:(length(intTest)-1), y = as.numeric(intTest)), color = "lightblue3") + 
  
  ggtitle("Original v.s. Interpolated Data") +
  
  labs(x = "time", 
       y = "value",
       label = "Legend Text")
  

    #error[k] <- sum(abs(orgTest-intTest))
}
  
  meanError[d] <- mean(error)
  D_vec[d] <- d

}

```

```{r}
plot(x = D,y = meanError, xlab = "Dataset index", ylab = "mean(sum(abs(x-X)))", type = "l")

p <- plot_ly(z = ~volcano) %>% add_surface(
  contours = list(
    z = list(
      show=TRUE,
      usecolormap=TRUE,
      highlightcolor="#ff0000",
      project=list(z=TRUE)
      )
    )
  ) %>%
  layout(
    scene = list(
      camera=list(
        eye = list(x=1.87, y=0.88, z=-0.64)
        )
      )
  )


x <- seq(0,100,length.out=100)
y <- seq(0,10,length.out=100)
z <- matrix(nrow = 100, ncol = 100)

for(i in 1:nrow(z)){
  for(j in 1:ncol(z)){
    z[i,j] = x[i]+y[j]
  }
}
p <- plot_ly(x=x,y=y,z=z) %>% add_surface()



### Creating a surface for Dataset D, x=p,y=g,z=MSE


MSEmat <- matrix(nrow=length(prop_vec),ncol=length(gap_vec))
rownames(MSEmat) <- prop_vec_names
colnames(MSEmat) <- gap_vec_names

z_mat <- list()
metric = "abs_differences"

for(d in 1:D){
  for(p in 1:P){
    for(g in 1:G){
      MSEmat[p,g] <- log(Evaluation[[d]][[p]][[g]][[m]][metric,"mean"])
   }
  }
  z_mat[[d]] <- MSEmat
  names(z_mat)[d] <- paste("D",d,sep="")
}


x <- prop_vec
y <- gap_vec
z <- MSEmat
plot_ly(data = data.frame(z_mat)) %>% 
  layout(title = paste(metric,collapse=""))
  add_surface(x=x,y=y,z=z_mat[[1]], opacity = 1*0.20, name = names(z_mat)[1]) %>%
  add_surface(x=x,y=y,z=z_mat[[2]], opacity = 2*0.20, name = names(z_mat)[2]) %>%
  add_surface(x=x,y=y,z=z_mat[[3]], opacity = 3*0.20, name = names(z_mat)[3]) %>%
  add_surface(x=x,y=y,z=z_mat[[4]], opacity = 4*0.20, name = names(z_mat)[4]) %>%
  add_surface(x=x,y=y,z=z_mat[[5]], opacity = 5*0.20, name = names(z_mat)[5]) 
  


```

