---
title: "Comparing the Performance of the HW Interpolator Against Other Known Interpolators"
author: "Sophie Castel"
date: "7/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

# Sandbox

## Preamble

```{r libraries, cache = TRUE}
library(multitaper)
library(tsinterp)
library(imputeTS)
library(zoo)
library(forecast)
library(MASS)
library(ggplot2)
library(dplyr)
library(snow)
library(parallel)
library(RColorBrewer)

# setting a seed for reproducibility
set.seed(997)
```


## Interpolators

Here we call some functions used for the interpolation of stationary time series.
```{r algorithms, cache = TRUE}
nearestNeighbor <- function(x) {
  stopifnot(is.ts(x)) 
  
  findNearestNeighbors <- function(x, i) {
    leftValid <- FALSE
    rightValid <- FALSE 
    numItLeft <- 1
    numItRight <- 1
    while (!leftValid) {
      leftNeighbor <- x[i - numItLeft]
      if (!is.na(leftNeighbor)) {
        leftValid <- TRUE
        leftNeighbor <- i - numItLeft
      }
      numItLeft <- numItLeft + 1
    }
    while (!rightValid) {
      rightNeighbor <- x[i + numItRight]
      if (!is.na(rightNeighbor)) {
        rightValid <- TRUE
        rightNeighbor <- i + numItRight
      }
      numItRight <- numItRight + 1
    }
    return(c(leftNeighbor, rightNeighbor))
  }
  
  for (i in 1:length(x)) {
    if (is.na(x[i])) {
      nearestNeighborsIndices <- findNearestNeighbors(x, i)
      a <- nearestNeighborsIndices[1]
      b <- nearestNeighborsIndices[2]
      if (i < ((a + b) / 2)) {
        x[i] <- x[a]
      } else {
        x[i] <- x[b]
      }
    }
  }
  return(x)
}

algorithm_names <- c("Nearest Neighbor",
                     "Linear Interpolation", 
                     "Natural Cubic Spline",
                     "FMM Cubic Spline", 
                     "Hermite Cubic Spline",
                     "Stineman Interpolation",
                     "Kalman - ARIMA",
                     "Kalman - StructTS",
                     "Last Observation Carried Forward",
                     "Next Observation Carried Backward", 
                     "Simple Moving Average", 
                     "Linear Weighted Moving Average",
                     "Exponential Weighted Moving Average",
                     "Replace with Mean",
                     "Replace with Median", 
                     "Replace with Mode",
                     "Replace with Random",
                     "Hybrid Wiener Interpolator")
algorithm_calls <- c("nearestNeighbor(", 
                     "na.approx(", 
                     "na.spline(method = 'natural', object = ",
                     "na.spline(method = 'fmm', object = ", 
                     "na.spline(method = 'monoH.FC', object = ", 
                     "na_interpolation(option = 'stine', x = ", 
                     "na_kalman(model = 'auto.arima', x = ", 
                     "na_kalman(model = 'StructTS', x = ",
                     "imputeTS::na.locf(option = 'locf', x = ", 
                     "imputeTS::na.locf(option = 'nocb', x = ", 
                     "na_ma(weighting = 'simple', x = ",
                     "na_ma(weighting = 'linear', x = ", 
                     "na_ma(weighting = 'exponential', x = ",
                     "na_mean(option = 'mean', x = ", 
                     "na_mean(option = 'median', x = ",
                     "na_mean(option = 'mode', x = ", 
                     "na_random(",
                     "interpolate(gap = which(is.na(x) == TRUE), progress = FALSE, z = ")

algorithms <- data.frame(algorithm_names, algorithm_calls)
```

# Interpolating the Gappy Data (in parallel)
Perform parallel interpolation on the Gappy Data using specified methods
```{r IntData, cache = TRUE}
methods <- c(18)
IntData <- list()

for(d in 1:length(OriginalData)){
  IntData[[d]] <- parInterpolate(data = GappyData[[d]], methods = methods)
}

names(IntData) <- names(OriginalData)
```

From this we get a multi-level list object called IntData, with dimension $d,m,p,g,k$:
$d$ = 1, ...,D = dataset$_d$
$m$ = 1, ...,M = interpolation method$_m$
$p$ = 1, ...,P length(prop_vec) = proportion of missing values$_p$
$g$ = 1, ...,G length(gap_vec) = gap width$_g$
$k$ = 1, ...,N = sample ID under each $d,m,p,g,k$ specification

# Computing the Performance Metrics
Applying the function to our Original and Interpolated Data:
```{r pmats}
pmats <- performance(x=OriginalData, X=IntData)
```

So, we have a list: $D \times M \times P \times G \times N \times C$ where $M$ is the number of interpolation methods used and $C=17$ is the number of performance criteria computed.

What are we interested in? The performance of each interpolation method with respect to the proportion missing and gap width. 
Thus, we want to compute the mean of each of the 17 performance criteria in each g,p,d,m specification across all k simulations.

# Evaluating the Performance Metrics
```{r Evaluation}
Evaluation <- evaluate(pmats = pmats, task = "mean")
```


